### 2021-04-27 doc/manuscript/Dow_manuscript_Nature.Rmd
update tree-ring figure
#49

#Figure numbers
schematic_summary=1 #illustrates parameters
pheno_Tsensitivity = 2 
DOYtiming = 3
  DOYtiming_SCBI = "3a"
  DOYtiming_Harvard = "3b"
tree_rings = 4  #stays, hopefully adding a lot more sites

@@

DOYtiming_caption <- "Figure 3 | Leaf (a,b) and stem growth (c,d) phenology at the Smithsonian Conservation Biology Institute (a,c) and Harvard Forest (b,d) from 1998 to 2020 (when available). Panels (a-b) show leaf phenology... Panels (c-d) show the dates at which stem growth milestones were acheived, on average, for sampled populations of ring-porous and diffuse-porous trees at each site. Mean temperature was calculated for each wood-type/site combination over the respective critical $T_{max}$ window, then turned into a ratio and assigned a color on a gradient where the coldest year in the sample is blue and the warmest is red. (**I THINK WE SHOULD COMBINE LEAF AND STEM PHENOLOGY ON A SINGLE PLOT WITH SEPARATE Y-AXES** See [GitHub Issue #50](https://github.com/EcoClimLab/growth_phenology/issues/50). )"

tree_rings_caption <- "Figure 4 | Sensitivity of annual growth, as derived from tree-rings, to monthly maximum temperatures, for ## chronologies from ## sites across eastern North America. Chronologies are grouped by xylem porosity and ordered by mean April T_max. Chronology details are given in Table S4. (NOTE: Figure still needs some work. See [GitHub Issue #49](https://github.com/EcoClimLab/growth_phenology/issues/49).)"

@@

```{r DOYtiming, out.width="100%", echo=FALSE, fig.align="center", fig.cap= DOYtiming_caption}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/DOYtiming_stemgrowth+foliage.png")
```

@@

We found that warmer spring temperatures in temperate deciduous forests cause earlier tree growth but have little effect on annual woody productivity ( `r schematic_summary`- `r DOYtiming`). Using dendroband data, we found that the peak growing season generally advanced as the critical spring windows warmed (Fig. ``r DOYtiming``). Despite this shift, growth variables $L_{pgs}$. $g_{max}$ and $\Delta DBH$ were virtually unchanged when this period warmed, but $DOY_{g_{max}}$ shifted back almost in unison with $DOY_{50}$ (Fig 2). The lack of influence Spring $T_{max}$ exudes on $\Delta DBH$ found with the dendroband data agrees with patterns observed in tree cores, which show that annual tree growth depends more on peak growing season conditions than early season advancement.

Although the number of years analyzed here is limited (10 years at SCBI, 6 years at HF), the signal is clear and consistent: for both species groups at both sites, warmer temperatures during virtually any part of the spring led to significantly earlier achievement of growth milestones (Figs. `r schematic_summary`-`r `r DOYtiming``).
This indicates that the effect of $T_{max}$ on the timing of growth ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$,and $DOY_{g_{max}}$) is not subtle. 
Indeed, growth milestones advanced by 1-8 days per degree Celsius during the critical temperature window.
This is consistent with previous studies showing that ... *(find/cite some statistics on how rapidly leaf phenology or start of C sequestration advance with T--- This could be tough. It looks like that varies alot by region and species.)*

@@

Our findings contribute to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (Fig. `r doytiming` , REFS), an earlier start to net C sequestration (REFS), and, as shown here, earlier initiation of stem growth (Figs. `r schematic_summary`- `r DOYtiming`).
Yet, an earlier start to the growing season is not associated with a longer growing season (Figs. `r schematic_summary`, `r pheno_Tsensitivity`) [@zani_increased_2020]
because growth finishes sooner under these conditions, presumably because of *C sink limitation* [@zani_increased_2020].

>>>>

#Figure numbers
schematic_summary=1 #illustrates parameters
pheno_Tsensitivity = 2 
DOYtiming = 3
  DOYtiming_SCBI = "3a"
  DOYtiming_Harvard = "3b"
tree_rings = 4

@@

DOYtiming_caption <- "Figure 3 | Leaf (a,b) and stem growth (c,d) phenology at the Smithsonian Conservation Biology Institute (a,c) and Harvard Forest (b,d) from 1998 to 2020 (when available). Panels (a-b) show leaf phenology... Panels (c-d) show the dates at which stem growth milestones were acheived, on average, for sampled populations of ring-porous and diffuse-porous trees at each site. Mean temperature was calculated for each wood-type/site combination over the respective critical $T_{max}$ window, then turned into a ratio and assigned a color on a gradient where the coldest year in the sample is blue and the warmest is red. (**SHOULD WE COMBINE LEAF AND STEM PHENOLOGY ON A SINGLE PLOT WITH SEPARATE Y-AXES?** See [GitHub Issue #50](https://github.com/EcoClimLab/growth_phenology/issues/50). )"

tree_rings_caption <- "Figure 4 | Sensitivity of annual growth, as derived from tree-rings, to monthly maximum temperatures, for ## chronologies from ## sites across eastern North America. Chronologies are grouped by xylem porosity and ordered by mean April T_max. Chronology details are given in Table S4. (NOTE: Figure still needs some work. **COLOR SCALE IS  A BIT OFF.**  See [GitHub Issue #49](https://github.com/EcoClimLab/growth_phenology/issues/49).)"

@@

```{r DOYtiming, out.width="100%", echo=FALSE, fig.align="center", fig.cap= DOYtiming_caption}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/DOYtiming_allyears.png")
```

@@

We found that warmer spring temperatures in temperate deciduous forests cause earlier tree growth but have little effect on annual woody productivity ( `r schematic_summary`- `r DOYtiming`). Using dendroband data, we found that the peak growing season generally advanced as the critical spring windows warmed (Fig. `r DOYtiming`). Despite this shift, growth variables $L_{pgs}$. $g_{max}$ and $\Delta DBH$ were virtually unchanged when this period warmed, but $DOY_{g_{max}}$ shifted back almost in unison with $DOY_{50}$ (Fig 2). The lack of influence Spring $T_{max}$ exudes on $\Delta DBH$ found with the dendroband data agrees with patterns observed in tree cores, which show that annual tree growth depends more on peak growing season conditions than early season advancement.

Although the number of years analyzed here is limited (10 years at SCBI, 6 years at HF), the signal is clear and consistent: for both species groups at both sites, warmer temperatures during virtually any part of the spring led to significantly earlier achievement of growth milestones (Figs. `r schematic_summary`-`r DOYtiming`).
This indicates that the effect of $T_{max}$ on the timing of growth ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$,and $DOY_{g_{max}}$) is not subtle. 
Indeed, growth milestones advanced by 1-8 days per degree Celsius during the critical temperature window.
This is consistent with previous studies showing that ... *(find/cite some statistics on how rapidly leaf phenology or start of C sequestration advance with T--- This could be tough. It looks like that varies alot by region and species.)*

@@

Our findings contribute to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (Fig. `r DOYtiming` , REFS), an earlier start to net C sequestration (REFS), and, as shown here, earlier initiation of stem growth (Figs. `r schematic_summary`- `r DOYtiming`).
Yet, an earlier start to the growing season is not associated with a longer growing season (Figs. `r schematic_summary`, `r pheno_Tsensitivity`) [@zani_increased_2020]
because growth finishes sooner under these conditions, presumably because of *C sink limitation* [@zani_increased_2020].
